---
title: "covid project 11/16/21 TS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#prework
rm(list=ls())
gc()
cat('\f')
library('fpp2')
library('dplyr')
library('janitor')
```
prework and load the packagge of fpp2


```{r}
setwd('C:/Users/nkheth1/Desktop/Time Series and Forecasting/Final Project')

#daily.covid.worldwide=read.csv('covid_19_data.csv') 
global.cases <- read.csv('time_series_covid_19_confirmed.csv')

#us.cases=read.csv('time_series_covid_19_confirmed_US.csv')
global.deaths <- read.csv('time_series_covid_19_deaths.csv')

#us.deaths=read.csv('time_series_covid_19_deaths_US.csv')  
global.recovered <- read.csv('time_series_covid_19_recovered.csv')
```
load the data under different names

Data Cleaning
```{r, echo=FALSE}
#df.list <- list(daily.covid.worldwide,global.cases,us.cases,global.deaths,us.deaths,global.recovered)
#lapply(df.list,anyNA
#daily.covid.worldwid=daily.covid.worldwide$SNo=NULL # just an ID variable, not useful for us

#confirmed cases data cleaning
global.cases <- global.cases[,-c(1,3,4)] #removing columns for region, latitude, and longitude
global.cases <- global.cases %>%  #sum different regional data under single country data
  group_by(Country.Region) %>% 
  summarise_all(sum)
global.cases.m <- as.matrix(global.cases)
global.cases.m.t <- t(global.cases.m) #convert data frame into matrix and transpose into columns
global.cases=as.data.frame(global.cases.m.t) #convert back into data frame
global.cases <- global.cases  %>% #use country (row 1) as column and variable names
  row_to_names(row_number(1))
global.cases <- sapply(global.cases, as.numeric) #convert data frame from string to numeric
anyNA(global.cases) #FALSE
cases.ts <- ts(global.cases[,c("Australia","China","Thailand", "US", "United Kingdom", "India")], 
               start = c(2020,23),
               frequency = 365)
cases.ts <- diff(cases.ts, 1)
global.cases <- as.data.frame(cases.ts)
write.csv(global.cases, "C://Users/nkheth1/Desktop/Time Series and Forecasting/Final Project/globalcases.csv",
          row.names = FALSE)


#confirmed deaths data cleaning
global.deaths <- global.deaths[,-c(1,3,4)]
global.deaths <- global.deaths %>% 
  group_by(Country.Region) %>% 
  summarise_all(sum)
global.deaths.m <- as.matrix(global.deaths)
global.deaths.m.t <- t(global.deaths.m)
global.deaths=as.data.frame(global.deaths.m.t)
global.deaths <- global.deaths  %>%
  row_to_names(row_number(1))
global.deaths <- sapply(global.deaths, as.numeric)
anyNA(global.deaths) #FALSE
deaths.ts <- ts(global.deaths[,c("Australia","China","Thailand", "US", "United Kingdom", "India")], 
                start = c(2020,23), 
                frequency = 365)
deaths.ts <- diff(deaths.ts, 1)
global.deaths <- as.data.frame(deaths.ts)
write.csv(global.deaths, "C://Users/nkheth1/Desktop/Time Series and Forecasting/Final Project/globaldeaths.csv",
          row.names = FALSE)

#confirmed recovered data cleaning
global.recovered <- global.recovered[,-c(1,3,4)]
global.recovered <- global.recovered %>% 
  group_by(Country.Region) %>% 
  summarise_all(sum)
global.recovered.m <- as.matrix(global.recovered)
global.recovered.m.t <- t(global.recovered.m)
global.recovered <- as.data.frame(global.recovered.m.t)
global.recovered <- global.recovered  %>%
  row_to_names(row_number(1))
global.recovered <- sapply(global.recovered, as.numeric)
anyNA(global.recovered) #FALSE
recovered.ts <- ts(global.recovered[,c("Australia","China","Thailand", "US", "United Kingdom", "India")], 
                   start = c(2020,23), 
                   frequency = 365)
recovered.ts <- diff(recovered.ts, 1)
global.recovered <- as.data.frame(recovered.ts)
write.csv(global.recovered, "C://Users/nkheth1/Desktop/Time Series and Forecasting/Final Project/globalrecovered.csv",
           row.names = FALSE)
```
While cleaning, saw that daily covid cases worldwide is formatted differently than a dataframe that can be converted to ts object. we could still convert it but we will have to switch up the format.has dates as a singular collumn not a collumn for each date as the others do. also some na's in the dataframes. we will need to reformate many of dataframes as they come as 
```{r, echo = FALSE}
###Time Series Elements and Partitioning

autoplot(cases.ts)
ggAcf(global.cases[,"Australia"], lag.max = 60) #trend
ggAcf(global.cases[,"China"], lag.max = 60) #trend
ggAcf(global.cases[,"Thailand"], lag.max = 60) #trend
ggAcf(global.cases[,"US"], lag.max = 60) #trend
ggAcf(global.cases[,"United Kingdom"], lag.max = 60) #trend
ggAcf(global.cases[,"India"], lag.max = 60) #trend

autoplot(deaths.ts)
ggAcf(global.deaths[,"Australia"], lag.max = 60) #trend
ggAcf(global.deaths[,"China"], lag.max = 60) #trend
ggAcf(global.deaths[,"Thailand"], lag.max = 60) #trend
ggAcf(global.deaths[,"US"], lag.max = 60) #trend, seems to have seasonality frequency 7 
ggAcf(global.deaths[,"United Kingdom"], lag.max = 60) #trend, also seem to have seasonality frequency 7
ggAcf(global.deaths[,"India"], lag.max = 60) #trend

autoplot(recovered.ts)
ggAcf(global.recovered[,"Australia"], lag.max = 60) #trend
ggAcf(global.recovered[,"China"], lag.max = 60) #trend
ggAcf(global.recovered[,"Thailand"], lag.max = 60) #trend
ggAcf(global.recovered[,"US"], lag.max = 60) #unreliable data
ggAcf(global.recovered[,"United Kingdom"], lag.max = 60) #trend
ggAcf(global.recovered[,"India"], lag.max = 60) #trend

cases.test <- tail(cases.ts, max(nrow(cases.ts)*0.2, 30))
cases.train <- head(cases.ts, nrow(cases.ts) - nrow(cases.test))

deaths.test <- tail(deaths.ts, max(nrow(cases.ts)*0.2, 30))
deaths.train <- head(deaths.ts, nrow(deaths.ts) - nrow(cases.test))

recovered.test <- tail(recovered.ts, max(nrow(recovered.ts)*0.2, 30))
recovered.train <- head(recovered.ts, nrow(recovered.ts) - nrow(recovered.test))
```


```{r, echo = FALSE}
### Model Building and Evaluation for Confirmed Cases

##Australia
cases.drift1 <- rwf(cases.train[,"Australia"], nrow(cases.test), drift = TRUE)
checkresiduals(cases.drift1) #not adequate
accuracy(cases.drift1, cases.test[,"Australia"])[2,] #RMSE = 11.825, MAPE = 74.03

cases.ses1 <- ses(cases.train[,"Australia"], nrow(cases.test))
checkresiduals(cases.ses1) #not adequate
accuracy(cases.ses1, cases.test[,"Australia"])[2,] #RMSE=10.892, MAPE = 63.101

cases.holt1 <- holt(cases.train[,"Australia"], nrow(cases.test))
checkresiduals(cases.holt1) #not adequate
accuracy(cases.holt1, cases.test[,"Australia"])[2,] #RMSE = 52.426, MAPE = 505.355

cases.arima1 <- auto.arima(cases.train[,"Australia"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
cases.arima1 #(p,d,q) == (1,1,3) AIC = 4076.03
checkresiduals(cases.arima1) #not adequate
cases.pred1 <- forecast(cases.arima1, nrow(cases.test))
accuracy(cases.pred1, cases.test[,"Australia"])[2,] #RMSE = 10.462, MAPE = 59.365

cases.arima1.1 <- auto.arima(cases.train[,"Australia"], seasonal = FALSE)
cases.arima1.1 #(p,d,q) == (1,1,3) AIC = 4076.03 (extremely high)
checkresiduals(cases.arima1.1) #not adequate
cases.pred1.1 <- forecast(cases.arima1.1, nrow(cases.test))
accuracy(cases.pred1.1, cases.test[,"Australia"])[2,] #RMSE = 10.462, MAPE = 59.365

##China
cases.drift2 <- rwf(cases.train[,"China"], nrow(cases.test), drift = TRUE)
checkresiduals(cases.drift2) #not adequate
accuracy(cases.drift2, cases.test[,"China"])[2,] #RMSE = 14.412, MAPE = 46.748

cases.ses2 <- ses(cases.train[,"China"], nrow(cases.test))
checkresiduals(cases.ses2) #not adequate
accuracy(cases.ses2, cases.test[,"China"])[2,] #RMSE= 9.94, MAPE = 36.582

cases.holt2 <- holt(cases.train[,"China"], nrow(cases.test))
checkresiduals(cases.holt2) #not adequate
accuracy(cases.holt2, cases.test[,"China"])[2,] #RMSE = 95.072, MAPE = 471.443

cases.arima2 <- auto.arima(cases.train[,"China"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
cases.arima2 #(p,d,q) == (0,1,2) AIC = 6357.25
checkresiduals(cases.arima2) #not adequate
cases.pred2 <- forecast(cases.arima2, nrow(cases.test))
accuracy(cases.pred2, cases.test[,"China"])[2,] #RMSE = 10.173, MAPE = 43.823

cases.arima2.1 <- auto.arima(cases.train[,"China"], seasonal = FALSE)
cases.arima2.1 #(p,d,q) == (2,1,1) AIC = 4076.03 (extremely high)
checkresiduals(cases.arima2.1) #not adequate
cases.pred2.1 <- forecast(cases.arima2.1, nrow(cases.test))
accuracy(cases.pred2.1, cases.test[,"China"])[2,] #RMSE = 9.936, MAPE = 37.203

##Thailand
cases.drift3 <- rwf(cases.train[,"Thailand"], nrow(cases.test), drift = TRUE)
checkresiduals(cases.drift3) #not adequate
accuracy(cases.drift3, cases.test[,"Thailand"])[2,] #RMSE = 1898.542, MAPE = Infinite

cases.ses3 <- ses(cases.train[,"Thailand"], nrow(cases.test))
checkresiduals(cases.ses3) #not adequate
accuracy(cases.ses3, cases.test[,"Thailand"])[2,] #RMSE= 1919.253, MAPE = Infinite

cases.holt3 <- holt(cases.train[,"Thailand"], nrow(cases.test))
checkresiduals(cases.holt3) #not adequate
accuracy(cases.holt3, cases.test[,"Thailand"])[2,] #RMSE = 1891.844, MAPE = Infinite

cases.arima3 <- auto.arima(cases.train[,"Thailand"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
cases.arima3 #(p,d,q) == (2,1,3) AIC = 4883.82
checkresiduals(cases.arima3) #not adequate
cases.pred3 <- forecast(cases.arima3, nrow(cases.test))
accuracy(cases.pred3, cases.test[,"Thailand"])[2,] #RMSE = 1843.526, MAPE = Infinite

#Rough search returns the same model as exhaustive search
cases.arima3.1 <- auto.arima(cases.train[,"Thailand"], seasonal = FALSE)
cases.arima3.1 #(p,d,q) == (2,1,3) AIC = 4883.32 (extremely high)
checkresiduals(cases.arima3.1) #not adequate
cases.pred3.1 <- forecast(cases.arima3.1, nrow(cases.test))
accuracy(cases.pred3.1, cases.test[,"Thailand"])[2,] #RMSE = 1843.526, MAPE = Infinite

##US
cases.drift4 <- rwf(cases.train[,"US"], nrow(cases.test), drift = TRUE)
checkresiduals(cases.drift4) #not adequate, does not eliminate trend from the data
accuracy(cases.drift4, cases.test[,"US"])[2,] #RMSE = 4.3x10^4, MAPE = 1.08x10^2

cases.ses4 <- ses(cases.train[,"US"], nrow(cases.test))
checkresiduals(cases.ses4) #not adequate, seasonality in the residuals
accuracy(cases.ses4, cases.test[,"US"])[2,] #RMSE= 2.75x10^4, MAPE = 68.8

cases.holt4 <- holt(cases.train[,"US"], nrow(cases.test))
checkresiduals(cases.holt4) #not adequate, seems to have some seasonality in residuals
accuracy(cases.holt4, cases.test[,"US"])[2,] #RMSE = 1.546x10^5, MAPE = 3.57x10^2

cases.arima4 <- auto.arima(cases.train[,"US"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
cases.arima4 #(p,d,q) == (3,1,2) AIC = 8700.13
checkresiduals(cases.arima4) #not adequate
cases.pred4 <- forecast(cases.arima4, nrow(cases.test))
accuracy(cases.pred4, cases.test[,"US"])[2,] #RMSE = 3.18x10^4, MAPE = 78.41

cases.arima4.1 <- auto.arima(cases.train[,"US"], seasonal = FALSE)
cases.arima4.1 #(p,d,q) == (0,1,1) AIC = 8756.75
checkresiduals(cases.arima4.1) #not adequate
cases.pred4.1 <- forecast(cases.arima4.1, nrow(cases.test))
accuracy(cases.pred4.1, cases.test[,"US"])[2,] #RMSE = 2.753x10^4, MAPE = 68.81

##United Kingdom
cases.drift5 <- rwf(cases.train[,"United Kingdom"], nrow(cases.test), drift = TRUE)
checkresiduals(cases.drift5) #not adequate, does not eliminate trend from the data
accuracy(cases.drift5, cases.test[,"United Kingdom"])[2,] #RMSE = 1.022x10^4, MAPE = 3.415x10^2

cases.ses5 <- ses(cases.train[,"United Kingdom"], nrow(cases.test))
checkresiduals(cases.ses5) #not adequate, seasonality in the residuals
accuracy(cases.ses5, cases.test[,"United Kingdom"])[2,] #RMSE= 8.536x10^3, MAPE = 2.836x10^2

cases.holt5 <- holt(cases.train[,"United Kingdom"], nrow(cases.test))
checkresiduals(cases.holt5) #not adequate, seems to have some seasonality in residuals
accuracy(cases.holt5, cases.test[,"United Kingdom"])[2,] #RMSE = 8.589x10^3, MAPE = 2.855x10^2

cases.arima5 <- auto.arima(cases.train[,"United Kingdom"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
cases.arima5 #(p,d,q) == (3,1,2) AIC = 7252.98
checkresiduals(cases.arima5) #not adequate
cases.pred5 <- forecast(cases.arima5, nrow(cases.test))
accuracy(cases.pred5, cases.test[,"United Kingdom"])[2,] #RMSE = 8.305x10^3, MAPE = 2.76x10^2

cases.arima5.1 <- auto.arima(cases.train[,"United Kingdom"], seasonal = FALSE)
cases.arima5.1 #(p,d,q) == (0,1,1) AIC = 7289.06
checkresiduals(cases.arima5.1) #not adequate
cases.pred5.1 <- forecast(cases.arima5.1, nrow(cases.test))
accuracy(cases.pred5.1, cases.test[,"United Kingdom"])[2,] #RMSE = 8.536x10^3, MAPE = 2.836x10^2

##India
cases.drift6 <- rwf(cases.train[,"India"], nrow(cases.test), drift = TRUE)
checkresiduals(cases.drift6) #not adequate, seems to have seasonality in the data
accuracy(cases.drift6, cases.test[,"India"])[2,] #RMSE = 2.072x10^5, MAPE = 71.08

cases.ses6 <- ses(cases.train[,"India"], nrow(cases.test))
checkresiduals(cases.ses6) #not adequate, seasonality in the residuals
accuracy(cases.ses6, cases.test[,"India"])[2,] #RMSE= 2.097x10^5, MAPE = 73.933

cases.holt6 <- holt(cases.train[,"India"], nrow(cases.test))
checkresiduals(cases.holt6) #not adequate, seems to have some seasonality in residuals
accuracy(cases.holt6, cases.test[,"India"])[2,] #RMSE = 2.023X10^5, MAPE = 6.886

cases.arima6 <- auto.arima(cases.train[,"India"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
cases.arima6 #(p,d,q) == (2,1,3) AIC = 7665.03
checkresiduals(cases.arima6) #not adequate, seasonality in the residuals
cases.pred6 <- forecast(cases.arima6, nrow(cases.test))
accuracy(cases.pred6, cases.test[,"India"])[2,] #RMSE = 2.076x10^5, MAPE = 71.73

#Rough search returns same model as exhaustive search 

cases.arima6.1 <- auto.arima(cases.train[,"India"], seasonal = FALSE)
cases.arima6.1 #(p,d,q) == (2,1,3) AIC = 7665/03
checkresiduals(cases.arima6.1) #not adequate
cases.pred6.1 <- forecast(cases.arima6.1, nrow(cases.test))
accuracy(cases.pred6.1, cases.test[,"India"])[2,] #RMSE = 2.076x10^5, MAPE = 71.73
```


```{r, echo = FALSE}
### Model Building and Evaluation for Deaths

##Australia
deaths.drift1 <- rwf(deaths.train[,"Australia"], nrow(deaths.test), drift = TRUE)
checkresiduals(deaths.drift1) #not adequate
accuracy(deaths.drift1, deaths.test[,"Australia"])[2,] #RMSE = 0.101, MAPE = 100.00

deaths.ses1 <- ses(deaths.train[,"Australia"], nrow(deaths.test))
checkresiduals(deaths.ses1) #not adequate
accuracy(deaths.ses1, deaths.test[,"Australia"])[2,] #RMSE= 0.101, MAPE = Infinite

deaths.holt1 <- holt(deaths.train[,"Australia"], nrow(deaths.test))
checkresiduals(deaths.holt1) #not adequate
accuracy(deaths.holt1, deaths.test[,"Australia"])[2,] #RMSE = 0.10, MAPE = Infinite

deaths.arima1 <- auto.arima(deaths.train[,"Australia"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
deaths.arima1 #(p,d,q) == (0,1,5) AIC = 2059.79
checkresiduals(deaths.arima1) #not adequate
deaths.pred1 <- forecast(deaths.arima1, nrow(deaths.test))
accuracy(deaths.pred1, deaths.test[,"Australia"])[2,] #RMSE = 0.101, MAPE = Infinite

#Rough search returns the same model as exhaustive search
deaths.arima1.1 <- auto.arima(deaths.train[,"Australia"], seasonal = FALSE)
deaths.arima1.1 #(p,d,q) == (3,1,2) AIC = 2076.63 (extremely high)
checkresiduals(deaths.arima1.1) #not adequate
deaths.pred1.1 <- forecast(deaths.arima1.1, nrow(deaths.test))
accuracy(deaths.pred1.1, deaths.test[,"Australia"])[2,] #RMSE = 0.101, MAPE = Infinite 
autoplot(deaths.ts[,"Australia"])

##China
autoplot(deaths.ts[,"China"])
deaths.drift2 <- rwf(deaths.train[,"China"], nrow(deaths.test), drift = TRUE)
checkresiduals(deaths.drift2) #not adequate
accuracy(deaths.drift2, deaths.test[,"China"])[2,] #RMSE = 0.441, MAPE = Infinite

deaths.ses2 <- ses(deaths.train[,"China"], nrow(deaths.test))
checkresiduals(deaths.ses2) #not adequate
accuracy(deaths.ses2, deaths.test[,"China"])[2,] #RMSE= 0.932, MAPE = Infinite

deaths.holt2 <- holt(deaths.train[,"China"], nrow(deaths.test))
checkresiduals(deaths.holt2) #not adequate
accuracy(deaths.holt2, deaths.test[,"China"])[2,] #RMSE = 3.303, MAPE = Infinite

deaths.arima2 <- auto.arima(deaths.train[,"China"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
deaths.arima2 #(p,d,q) == (0,1,1) AIC = 4446.8
checkresiduals(deaths.arima2) #not adequate
deaths.pred2 <- forecast(deaths.arima2, nrow(deaths.test))
accuracy(deaths.pred2, deaths.test[,"China"])[2,] #RMSE = 0.39, MAPE = Infinite

#Rough search returns the same model as exhaustive search
deaths.arima2.1 <- auto.arima(deaths.train[,"China"], seasonal = FALSE)
deaths.arima2.1 #(p,d,q) == (1,1,2) AIC = 4449.21
checkresiduals(deaths.arima2.1) #not adequate
deaths.pred2.1 <- forecast(deaths.arima2.1, nrow(deaths.test))
accuracy(deaths.pred2.1, deaths.test[,"China"])[2,] #RMSE = 0.912, MAPE = Infinite

##Thailand
autoplot(deaths.ts[,"Thailand"])
deaths.drift3 <- rwf(deaths.train[,"Thailand"], nrow(deaths.test), drift = TRUE)
checkresiduals(deaths.drift3) #not adequate
accuracy(deaths.drift3, deaths.test[,"Thailand"])[2,] #RMSE = 15.574, MAPE = 100

deaths.ses3 <- ses(deaths.train[,"Thailand"], nrow(deaths.test))
checkresiduals(deaths.ses3) #not adequate
accuracy(deaths.ses3, deaths.test[,"Thailand"])[2,] #RMSE= 15.361, MAPE = Infinite

deaths.holt3 <- holt(deaths.train[,"Thailand"], nrow(deaths.test))
checkresiduals(deaths.holt3) #not adequate
accuracy(deaths.holt3, deaths.test[,"Thailand"])[2,] #RMSE = 15.313, MAPE = Infinite

deaths.arima3 <- auto.arima(deaths.train[,"Thailand"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
deaths.arima3 #(p,d,q) == (2,1,1) AIC = 497.44
checkresiduals(deaths.arima3) #not adequate
deaths.pred3 <- forecast(deaths.arima3, nrow(deaths.test))
accuracy(deaths.pred3, deaths.test[,"Thailand"])[2,] #RMSE = 15.362, MAPE = Infinite

#Rough search returns the same model as exhaustive search
deaths.arima3.1 <- auto.arima(deaths.train[,"Thailand"], seasonal = FALSE)
deaths.arima3.1 #(p,d,q) == (2,1,1) AIC = 497.44
checkresiduals(deaths.arima3.1) #not adequate
deaths.pred3.1 <- forecast(deaths.arima3.1, nrow(deaths.test))
accuracy(deaths.pred3.1, deaths.test[,"Thailand"])[2,] #RMSE = 15.362, MAPE = Infinite

##US
deaths.drift4 <- rwf(deaths.train[,"US"], nrow(deaths.test), drift = TRUE)
checkresiduals(deaths.drift4) #not adequate, there is some seasonality in the residuals
accuracy(deaths.drift4, deaths.test[,"US"])[2,] #RMSE = 2.116x10^3, MAPE = 3.276x10^2

deaths.ses4 <- ses(deaths.train[,"US"], nrow(deaths.test))
checkresiduals(deaths.ses4) #not adequate, there is some seasonality in the residuals
accuracy(deaths.ses4, deaths.test[,"US"])[2,] #RMSE= 1.754x10^3, MAPE = 2.714x10^2

deaths.holt4 <- holt(deaths.train[,"US"], nrow(deaths.test))
checkresiduals(deaths.holt4) #not adequate, still some seasonality left 
accuracy(deaths.holt4, deaths.test[,"US"])[2,] #RMSE = 1.761x10^3, MAPE = 2.726x10^2

deaths.arima4 <- auto.arima(deaths.train[,"US"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
deaths.arima4 #(p,d,q) == (5,1,0) AIC = 5605.79
checkresiduals(deaths.arima4) #not adequate
deaths.pred4 <- forecast(deaths.arima4, nrow(deaths.test))
accuracy(deaths.pred4, deaths.test[,"US"])[2,] #RMSE = 1.144x10^3, MAPE = 1.79x10^2

#Rough search returns the same model as exhaustive search
deaths.arima4.1 <- auto.arima(deaths.train[,"US"], seasonal = FALSE)
deaths.arima4.1 #(p,d,q) == (3,1,2) AIC = 5664.33
checkresiduals(deaths.arima4.1) #not adequate
deaths.pred4.1 <- forecast(deaths.arima4.1, nrow(deaths.test))
accuracy(deaths.pred4.1, deaths.test[,"US"])[2,] #RMSE = 1.032x10^3, MAPE = 1.616x10^2

##Unite Kingdom
deaths.drift5 <- rwf(deaths.train[,"United Kingdom"], nrow(deaths.test), drift = TRUE)
checkresiduals(deaths.drift5) #not adequate, there is some seasonality in the residuals
accuracy(deaths.drift5, deaths.test[,"United Kingdom"])[2,] #RMSE = 5.41x10^2, MAPE = 4.655x10^3

deaths.ses5 <- ses(deaths.train[,"United Kingdom"], nrow(deaths.test))
checkresiduals(deaths.ses5) #not adequate, there is some seasonality in the residuals
accuracy(deaths.ses5, deaths.test[,"United Kingdom"])[2,] #RMSE= 4.974x10^2, MAPE = 4.134x10^3

deaths.holt5 <- holt(deaths.train[,"United Kingdom"], nrow(deaths.test))
checkresiduals(deaths.holt5) #not adequate, still some seasonality left 
accuracy(deaths.holt5, deaths.test[,"United Kingdom"])[2,] #RMSE = 1.689x10^2, MAPE = 1.402x10^3

deaths.arima5 <- auto.arima(deaths.train[,"United Kingdom"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
deaths.arima5 #(p,d,q) == (5,1,0) AIC = 4944.39
checkresiduals(deaths.arima5) #not adequate
deaths.pred5 <- forecast(deaths.arima5, nrow(deaths.test))
accuracy(deaths.pred5, deaths.test[,"United Kingdom"])[2,] #RMSE = 4.532x10^2, MAPE = 3.803x10^3

#Rough search returns the same model as exhaustive search
deaths.arima5.1 <- auto.arima(deaths.train[,"United Kingdom"], seasonal = FALSE)
deaths.arima5.1 #(p,d,q) == (1,1,5) AIC = 5003.89
checkresiduals(deaths.arima5.1) #not adequate
deaths.pred5.1 <- forecast(deaths.arima5.1, nrow(deaths.test))
accuracy(deaths.pred5.1, deaths.test[,"United Kingdom"])[2,] #RMSE = 4.215x10^2, MAPE = 3.548x10^3

##India
deaths.drift6 <- rwf(deaths.train[,"India"], nrow(deaths.test), drift = TRUE)
checkresiduals(deaths.drift6) #not adequate
accuracy(deaths.drift6, deaths.test[,"India"])[2,] #RMSE = 2.303x10^3, MAPE = 67.7

deaths.ses6 <- ses(deaths.train[,"India"], nrow(deaths.test))
checkresiduals(deaths.ses6) #not adequate
accuracy(deaths.ses6, deaths.test[,"India"])[2,] #RMSE= 2.32x10^3, MAPE = 69.573

deaths.holt6 <- holt(deaths.train[,"India"], nrow(deaths.test))
checkresiduals(deaths.holt6) #not adequate 
accuracy(deaths.holt6, deaths.test[,"India"])[2,] #RMSE = 2.452x10^3, MAPE = 87.9

deaths.arima6 <- auto.arima(deaths.train[,"India"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
deaths.arima6 #(p,d,q) == (1,1,3) AIC = 4830.06
checkresiduals(deaths.arima6) #not adequate
deaths.pred6 <- forecast(deaths.arima6, nrow(deaths.test))
accuracy(deaths.pred6, deaths.test[,"India"])[2,] #RMSE = 2.321x10^3, MAPE = 70.153

#Rough search returns the same model as exhaustive search
deaths.arima6.1 <- auto.arima(deaths.train[,"India"], seasonal = FALSE)
deaths.arima6.1 #(p,d,q) == (1,1,3) AIC = 5664.33
checkresiduals(deaths.arima6.1) #not adequate
deaths.pred6.1 <- forecast(deaths.arima6.1, nrow(deaths.test))
accuracy(deaths.pred6.1, deaths.test[,"India"])[2,] #RMSE = 2.321x10^3, MAPE = 70.153

```


```{r, echo = FALSE}
### Model Building and Evaluation for Recovered

##Australia
recovered.drift1 <- rwf(recovered.train[,"Australia"], nrow(recovered.test), drift = TRUE)
checkresiduals(recovered.drift1) #not adequate
accuracy(recovered.drift1, recovered.test[,"Australia"])[2,] #RMSE = 7.175, MAPE = Infinite

recovered.ses1 <- ses(recovered.train[,"Australia"], nrow(recovered.test))
checkresiduals(recovered.ses1) #not adequate
accuracy(recovered.ses1, recovered.test[,"Australia"])[2,] #RMSE= 7.767, MAPE = Infinite

recovered.holt1 <- holt(recovered.train[,"Australia"], nrow(recovered.test))
checkresiduals(recovered.holt1) #not adequate
accuracy(recovered.holt1, recovered.test[,"Australia"])[2,] #RMSE = 6.945, MAPE = Infinite

###best MAPE
recovered.arima1 <- auto.arima(recovered.train[,"Australia"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
recovered.arima1 #(p,d,q) == (1,0,2) AIC = 5364.41
checkresiduals(recovered.arima1) #not adequate
recovered.pred1 <- forecast(recovered.arima1, nrow(recovered.test))
accuracy(recovered.pred1, recovered.test[,"Australia"])[2,] #RMSE = 44.965, MAPE = Infinite

#Rough search returns the same model as exhaustive search
recovered.arima1.1 <- auto.arima(recovered.train[,"Australia"], seasonal = FALSE)
recovered.arima1.1 #(p,d,q) == (1,0,2) AIC = 5364.41 (extremely high)
checkresiduals(recovered.arima1.1) #not adequate
recovered.pred1.1 <- forecast(recovered.arima1.1, nrow(recovered.test))
accuracy(recovered.pred1.1, recovered.test[,"Australia"])[2,] #RMSE = 44.965, MAPE = Infinite 

##China
autoplot(recovered.ts[,"China"])
recovered.drift2 <- rwf(recovered.train[,"China"], nrow(recovered.test), drift = TRUE)
checkresiduals(recovered.drift2) #not adequate
accuracy(recovered.drift2, recovered.test[,"China"])[2,] #RMSE = 61.791, MAPE = 307.04

recovered.ses2 <- ses(recovered.train[,"China"], nrow(recovered.test))
checkresiduals(recovered.ses2) #not adequate
accuracy(recovered.ses2, recovered.test[,"China"])[2,] #RMSE= 54.61, MAPE = 270.254

recovered.holt2 <- holt(recovered.train[,"China"], nrow(recovered.test))
checkresiduals(recovered.holt2) #not adequate
accuracy(recovered.holt2, recovered.test[,"China"])[2,] #RMSE = 351.246, MAPE = 1.61x10^3

recovered.arima2 <- auto.arima(recovered.train[,"China"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
recovered.arima2 #(p,d,q) == (0,1,2) AIC = 5296.13
checkresiduals(recovered.arima2) #not adequate
recovered.pred2 <- forecast(recovered.arima2, nrow(recovered.test))
accuracy(recovered.pred2, recovered.test[,"China"])[2,] #RMSE = 55.401, MAPE = 274.091

#Rough search returns the same model as exhaustive search
recovered.arima2.1 <- auto.arima(recovered.train[,"China"], seasonal = FALSE)
recovered.arima2.1 #(p,d,q) == (0,1,2) AIC = 5296.13
checkresiduals(recovered.arima2.1) #not adequate
recovered.pred2.1 <- forecast(recovered.arima2.1, nrow(recovered.test))
accuracy(recovered.pred2.1, recovered.test[,"China"])[2,] #RMSE = 55.401, MAPE = 274.091

##Thailand
recovered.drift3 <- rwf(recovered.train[,"Thailand"], nrow(recovered.test), drift = TRUE)
checkresiduals(recovered.drift3) #not adequate
accuracy(recovered.drift3, recovered.test[,"Thailand"])[2,] #RMSE = 70.669, MAPE = Infinite

recovered.ses3 <- ses(recovered.train[,"Thailand"], nrow(recovered.test))
checkresiduals(recovered.ses3) #not adequate
accuracy(recovered.ses3, recovered.test[,"Thailand"])[2,] #RMSE= 435.89, MAPE = Infinite

recovered.holt3 <- holt(recovered.train[,"Thailand"], nrow(recovered.test))
checkresiduals(recovered.holt3) #not adequate
accuracy(recovered.holt3, recovered.test[,"Thailand"])[2,] #RMSE = 511.855, MAPE = Infinite

recovered.arima3 <- auto.arima(recovered.train[,"Thailand"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
recovered.arima3 #(p,d,q) == (0,1,5) AIC = 4956.9
checkresiduals(recovered.arima3) #not adequate
recovered.pred3 <- forecast(recovered.arima3, nrow(recovered.test))
accuracy(recovered.pred3, recovered.test[,"Thailand"])[2,] #RMSE = 312.458, MAPE = Infinite

#Rough search returns the same model as exhaustive search
recovered.arima3.1 <- auto.arima(recovered.train[,"Thailand"], seasonal = FALSE)
recovered.arima3.1 #(p,d,q) == (2,1,2) AIC = 4966.08
checkresiduals(recovered.arima3.1) #not adequate
recovered.pred3.1 <- forecast(recovered.arima3.1, nrow(recovered.test))
accuracy(recovered.pred3.1, recovered.test[,"Thailand"])[2,] #RMSE = 303.01, MAPE = Infinite

##US
#recovered.drift4 <- rwf(recovered.train[,"US"], nrow(recovered.test), drift = TRUE)
#checkresiduals(recovered.drift4) #not adequate, there is some seasonality in the residuals
#accuracy(recovered.drift4, recovered.test[,"US"])[2,] #RMSE = 2.116x10^3, MAPE = 3.276x10^2

#recovered.ses4 <- ses(recovered.train[,"US"], nrow(recovered.test))
#checkresiduals(recovered.ses4) #not adequate, there is some seasonality in the residuals
#accuracy(recovered.ses4, recovered.test[,"US"])[2,] #RMSE= 1.754x10^3, MAPE = 2.714x10^2

#recovered.holt4 <- holt(recovered.train[,"US"], nrow(recovered.test))
#checkresiduals(recovered.holt4) #not adequate, still some seasonality left 
#accuracy(recovered.holt4, recovered.test[,"US"])[2,] #RMSE = 1.761x10^3, MAPE = 2.726x10^2

#recovered.arima4 <- auto.arima(recovered.train[,"US"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
#recovered.arima4 #(p,d,q) == (5,1,0) AIC = 5605.79
#checkresiduals(recovered.arima4) #not adequate
#recovered.pred4 <- forecast(recovered.arima4, nrow(recovered.test))
#accuracy(recovered.pred4, recovered.test[,"US"])[2,] #RMSE = 1.144x10^3, MAPE = 1.79x10^2

#Rough search returns the same model as exhaustive search
#recovered.arima4.1 <- auto.arima(recovered.train[,"US"], seasonal = FALSE)
#recovered.arima4.1 #(p,d,q) == (3,1,2) AIC = 5664.33
#checkresiduals(recovered.arima4.1) #not adequate
#recovered.pred4.1 <- forecast(recovered.arima4.1, nrow(recovered.test))
#accuracy(recovered.pred4.1, recovered.test[,"US"])[2,] #RMSE = 1.032x10^3, MAPE = 1.616x10^2

##Unite Kingdom
recovered.drift5 <- rwf(recovered.train[,"United Kingdom"], nrow(recovered.test), drift = TRUE)
checkresiduals(recovered.drift5) #not adequate, there is some seasonality in the residuals
accuracy(recovered.drift5, recovered.test[,"United Kingdom"])[2,] #RMSE = 72.28, MAPE = Infinite

recovered.ses5 <- ses(recovered.train[,"United Kingdom"], nrow(recovered.test))
checkresiduals(recovered.ses5) #not adequate, there is some seasonality in the residuals
accuracy(recovered.ses5, recovered.test[,"United Kingdom"])[2,] #RMSE= 58.287, MAPE = Infinite

recovered.holt5 <- holt(recovered.train[,"United Kingdom"], nrow(recovered.test))
checkresiduals(recovered.holt5) #not adequate, still some seasonality left 
accuracy(recovered.holt5, recovered.test[,"United Kingdom"])[2,] #RMSE = 68.894, MAPE = Infinite

recovered.arima5 <- auto.arima(recovered.train[,"United Kingdom"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
recovered.arima5 #(p,d,q) == (1,1,4) AIC = 4019.25
checkresiduals(recovered.arima5) #not adequate
recovered.pred5 <- forecast(recovered.arima5, nrow(recovered.test))
accuracy(recovered.pred5, recovered.test[,"United Kingdom"])[2,] #RMSE = 57.819, MAPE = Infinite

#Rough search returns the same model as exhaustive search
recovered.arima5.1 <- auto.arima(recovered.train[,"United Kingdom"], seasonal = FALSE)
recovered.arima5.1 #(p,d,q) == (1,1,4) AIC = 4019.03
checkresiduals(recovered.arima5.1) #not adequate
recovered.pred5.1 <- forecast(recovered.arima5.1, nrow(recovered.test))
accuracy(recovered.pred5.1, recovered.test[,"United Kingdom"])[2,] #RMSE = 57.819, MAPE = Infinite

##India
recovered.drift6 <- rwf(recovered.train[,"India"], nrow(recovered.test), drift = TRUE)
checkresiduals(recovered.drift6) #not adequate
accuracy(recovered.drift6, recovered.test[,"India"])[2,] #RMSE = 1.956x10^5, MAPE = 70.242

recovered.ses6 <- ses(recovered.train[,"India"], nrow(recovered.test))
checkresiduals(recovered.ses6) #not adequate
accuracy(recovered.ses6, recovered.test[,"India"])[2,] #RMSE= 1.965x10^5, MAPE = 70.237

recovered.holt6 <- holt(recovered.train[,"India"], nrow(recovered.test))
checkresiduals(recovered.holt6) #not adequate 
accuracy(recovered.holt6, recovered.test[,"India"])[2,] #RMSE = 2.086x10^5, MAPE = 84.51

recovered.arima6 <- auto.arima(recovered.train[,"India"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
recovered.arima6 #(p,d,q) == (2,2,3) AIC = 7632.19
checkresiduals(recovered.arima6) #not adequate
recovered.pred6 <- forecast(recovered.arima6, nrow(recovered.test))
accuracy(recovered.pred6, recovered.test[,"India"])[2,] #RMSE = 2.082x10^5, MAPE = 83.82

#Rough search returns the same model as exhaustive search
recovered.arima6.1 <- auto.arima(recovered.train[,"India"], seasonal = FALSE)
recovered.arima6.1 #(p,d,q) == (0,2,2) AIC = 7636.97
checkresiduals(recovered.arima6.1) #not adequate
recovered.pred6.1 <- forecast(recovered.arima6.1, nrow(recovered.test))
accuracy(recovered.pred6.1, recovered.test[,"India"])[2,] #RMSE = 2.086x10^5, MAPE = 84.547
```

Forecasts for Confirmed Cases
```{r, echo = FALSE}

#Australia
cases.arima1 <- Arima(cases.ts[,"Australia"], order = c(1,1,3))
cases.pred1 <- forecast(cases.arima1, 30)
autoplot(cases.pred1) + autolayer(fitted(cases.pred1), series = "Fitted")

#China
cases.arima2 <- Arima(cases.ts[,"China"], order = c(2,1,1))
cases.pred2 <- forecast(cases.arima2, 30)
autoplot(cases.pred2) + autolayer(fitted(cases.pred2), series = "Fitted")

#Thailand
cases.arima3 <- Arima(cases.ts[,"Thailand"], order = c(2,1,3))
cases.pred3 <- forecast(cases.arima3, 30)
autoplot(cases.pred3) + autolayer(fitted(cases.pred3), series = "Fitted")

#US
cases.arima4 <- Arima(cases.ts[,"US"], order = c(0,1,1))
cases.pred4 <- forecast(cases.arima4, 30)
autoplot(cases.pred4) + autolayer(fitted(cases.pred4), series = "Fitted")

#United Kingdom
cases.arima5 <- Arima(cases.ts[,"United Kingdom"], order = c(3,1,2))
cases.pred5 <- forecast(cases.arima5, 30)
autoplot(cases.pred5) + autolayer(fitted(cases.pred5), series = "Fitted")

#India
cases.arima6 <- Arima(cases.ts[,"India"], order = c(2,1,3))
cases.pred6 <- forecast(cases.arima6, 30)
autoplot(cases.pred6) + autolayer(fitted(cases.pred6), series = "Fitted")

```


```{r, echo = FALSE}
#Australia
deaths.arima1 <- Arima(deaths.ts[,"Australia"], order = c(0,1,5))
deaths.pred1 <- forecast(deaths.arima1, 30)
autoplot(deaths.pred1) + autolayer(fitted(deaths.pred1), series = "Fitted")

#China
deaths.arima2 <- Arima(deaths.ts[,"China"], order = c(0,1,1))
deaths.pred2 <- forecast(deaths.arima2, 30)
autoplot(deaths.pred2) + autolayer(fitted(deaths.pred2), series = "Fitted")

#Thailand
deaths.arima3 <- Arima(deaths.ts[,"Thailand"], order = c(2,1,1))
deaths.pred3 <- forecast(deaths.arima3, 30)
autoplot(deaths.pred3) + autolayer(fitted(deaths.pred3), series = "Fitted")

#US
deaths.arima4 <- Arima(deaths.ts[,"US"], order = c(3,1,2))
deaths.pred4 <- forecast(deaths.arima4, 30)
autoplot(deaths.pred4) + autolayer(fitted(deaths.pred4), series = "Fitted")

#United Kingdom
deaths.arima5 <- Arima(deaths.ts[,"United Kingdom"], order = c(1,1,5))
deaths.pred5 <- forecast(deaths.arima5, 30)
autoplot(deaths.pred5) + autolayer(fitted(deaths.pred5), series = "Fitted")

#India
deaths.arima6 <- Arima(deaths.ts[,"India"], order = c(1,1,3))
deaths.pred6 <- forecast(deaths.arima6, 30)
autoplot(deaths.pred6) + autolayer(fitted(deaths.pred6), series = "Fitted")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
